---
title: "Climate change evidence"
output:
  pdf_document: 
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
# Set libraries
library(readxl)
library(ggplot2)
library(gdata)
library(tidyverse)
library(esquisse)
library(knitr)
library(zoo)
library(psych)

# Define global variables
my_theme <- theme_classic() + 
            theme(text = element_text(family = "serif"),
                  plot.title = element_text(size = 20, face = "bold"),
                                            #margin = margin(t = 6, r = 0, b = 4, l = 0, unit = "pt")),
                  plot.subtitle = element_text(size = 14),
                  plot.caption = element_text(hjust = 0),
                  axis.title.y = element_text(size = 11, hjust = 1),
                  axis.title.x = element_text(size = 11, hjust = 0),
                  legend.position = "bottom"
                ) 

# Option to see esquisse
options("esquisse.display.mode" = "dialog")
```

# Data processing

## NASA/GISS data
Analyze global land-ocean data. No need of pre-processing, data already clean.

```{r}
df_temp_anomaly <- read.csv("data/NASA/global_temperature_anomaly.csv", 
                            sep = ",", header = TRUE, skip = 2)

# Have a look at first 5 row of the dataset
kable(head(df_temp_anomaly))
```


## Berkley Earth data 
Have a look at different data per continent.

### Analysis of one continent file

Do the processing of one single file, i.e. data processing of one continent to have a better idea of what to do
```{r}
# Define columns data frame
cols_month <- c("year", "month", "monthly_anomaly", 
                "monthly_uncertainty", "annual_anomaly", 
                "annual_uncertainty", "five_year_anomaly", 
                "five_year_uncertainty", "ten_year_anomaly", 
                "ten_year_uncertainty", "twenty_year_anomaly", "twenty_year_uncertainty")

path <- "data/Berkley Earth/continent_monthly_avg_temp/europe-TAVG-Trend.txt"

# Read data
df_europe_month <- read_delim(path,
                        delim = " ",
                        col_names = cols_month,
                        # Specify all columns as character otherwise error show up
                        col_types = "cccccccccccc",
                        skip = 70, 
                        na = c("", "NA","NaN"))
  
# Only keep first 4 columns
df_europe_month <- df_europe_month[, 1:4]

# Convert all columns to numeric 
df_europe_month <- data.frame(sapply(df_europe_month, as.numeric))

# Get name of zone from the path name
zone_name_path <- unlist(strsplit(path, split = "/"))[4]

# Get name of the zone as a list
zone_name <- head(unlist(strsplit(zone_name_path, split = "-")), -2)

# Collapse name into a single string
zone_name <- paste(zone_name, collapse = '_')

# Add name of the zone as a column to the dataset
df_europe_month$zone <- zone_name

# Create a single date column
df_europe_month$date = as.yearmon(paste(df_europe_month$year, df_europe_month$month), "%Y %m")

# Delete column year and month
df_europe_month <- subset(df_europe_month, select = -c(year,month))
  
dim(df_europe_month)
str(df_europe_month)
kable(head(df_europe_month))
kable(summary(df_europe_month))
```

### Aggregate all data 

Using a for loop, read all the files per single continent contained in the same folder and apply the necessary cleaning

```{r}
# Get a list with all the files inside the folder zone_monthly_average_temp
files <- list.files(path="data/Berkley Earth/continent_monthly_avg_temp", 
                    pattern="*.txt", 
                    full.names=TRUE, 
                    recursive=FALSE)

# Create empty data frame with correct columns to store all the data
df_continent_avg_temp <- data.frame(zone = character(),
                                   year = character(),
                                   anomaly = double(),
                                   lowess = double(),
                                   uncertainty = double(),
                                   stringsAsFactors=FALSE)

# Loop true all the files in the folder and read them
for (file in files){
  
  # Read and create a temporary dataset
  df_year_temp <- read_delim(file,
                        delim = " ",
                        col_names = cols_month,
                        # Specify all columns as character otherwise error show up
                        col_types = "cccccccccccc",
                        skip = 70, 
                        na = c("", "NA","NaN"))
  
  # Only keep first 4 columns
  df_year_temp <- df_year_temp[, 1:4]
  
  # Convert all columns to numeric 
  df_year_temp <- data.frame(sapply(df_year_temp, as.numeric))
  
  # CALCULATE AVERAGE PER YEAR GIVEN MONTHLY DATA

  # Aggregate data per year and calculate mean
  df_year_temp <- aggregate(monthly_anomaly~year, transform(df_year_temp, year = df_year_temp$year), mean)
  
  # Change name of the column 
  colnames(df_year_temp)[2] <- "yearly_anomaly"

  # Compute LOWESS smoother which uses locally-weighted polynomial regression
  line_trend <- lowess(as.numeric(df_year_temp$year), df_year_temp$yearly_anomaly, f = 1/5)
  
  # Add lowess to year temp dataframe as new column
  df_year_temp$lowess <- line_trend$y
  
  # GET ZONE NAME FOR PATH
  
  # Unlist path name
  zone_name_path <- unlist(strsplit(file, split = "/"))[4]
  
  # Get name of the zone as a list
  zone_name <- head(unlist(strsplit(zone_name_path, split = "-")), -2)
  
  # Collapse name into a single string
  zone_name <- paste(zone_name, collapse = '_')
  
  # Add name of the zone as a column to the temporary dataset
  df_year_temp$zone <- zone_name
  
  # Update dataset
  df_continent_avg_temp <- rbind(df_continent_avg_temp, df_year_temp)
}
```

# Descriptive Statistics

## NASA/GISS

```{r}
# Get dimension of the database
dim(df_temp_anomaly)

# Get information about columns
str(df_temp_anomaly)

# Get summary information about columns
kable(summary(df_temp_anomaly))

# Get summary information about columns
kable(summary(df_temp_anomaly))
```

## Berkley Earth

```{r}
dim(df_continent_avg_temp)
kable(summary(df_continent_avg_temp))
str(df_continent_avg_temp)
kable(head(df_continent_avg_temp))
kable(describe(df_continent_avg_temp)[1:7])
kable(describe(df_continent_avg_temp)[7:13])
```

### Descriptitive statistics per continent

Using describeBy() function of psych package.

```{r}
describeBy(df_continent_avg_temp, df_continent_avg_temp$zone)
```

Manual approach.

```{r}
# Name of continents
continents <- c("africa", "europe", "asia", "oceania", 
                "north_america", "south_america", "antarctica")

# Create empty vectors
min_year <- c()
min_temp_anom <- c()
max_temp_anom <- c()
mean_temp_anom <- c()
var_temp_anom <- c()

# Loop through all the continents
for (continent in continents){
  
  # Filter the continent
  df_continent <- df_continent_avg_temp[df_continent_avg_temp$zone == continent, ]
  
  # Get descriptive statistics per continent
  min_year <- c(min_year, min(df_continent$year))
  min_temp_anom <- c(min_temp_anom, min(df_continent$yearly_anomaly))
  max_temp_anom <- c(max_temp_anom, max(df_continent$yearly_anomaly))
  mean_temp_anom <- c(mean_temp_anom, mean(df_continent$yearly_anomaly))
  var_temp_anom <- c(var_temp_anom, var(df_continent$yearly_anomaly))
}

df_summary_continent <- data.frame(continents, min_year, min_temp_anom, 
                                   max_temp_anom, mean_temp_anom, var_temp_anom)

kable(df_summary_continent)
```

# Data Visualization

## Global temperature

```{r, echo=F, fig.width = 9, fig.height = 6}

cols <- c("red", "black")



ggplot(df_temp_anomaly, aes(x = Year)) +
  # Add horizontal and vertical line
  geom_hline(yintercept = 0, linetype="dashed", color = "grey") +
  geom_vline(xintercept = 1951, linetype="dashed", color = "grey") +
  geom_vline(xintercept = 1980, linetype="dashed", color = "grey") +
  # Add global annual mean point connected by a line
  geom_point(aes(y = No_Smoothing, colour = "No_Smoothing")) +
  geom_line(aes(y = No_Smoothing, colour = "No_Smoothing"), size = 0.5) +
  # Add five-year lowess smooth line
  geom_line(aes(y = Lowess.5., colour = "Lowess"), size = 0.5) +
  # Personalize legend
  guides(color = guide_legend(title = "")) +
  # Manual color 
  scale_color_manual(values = cols, labels = c("Lowess Smoothing", "Annual Mean")) + 
  # Add more ticks to x axis
  scale_x_continuous(breaks = round(seq(min(df_temp_anomaly$Year), max(df_temp_anomaly$Year), by = 10), 1),
                     expand = c(0.007, 0)) + 
  annotate("text", x=1997, y=0.9, label= "Nineteen of the hottest years 
have occurred since 2000", size = 2.5) +
  # Add labels and text
  labs(
    x = "Year",
    y = "Temperature anomaly w.r.t 1951-80 (°C)",
    title = "Global Temperature",
    subtitle = "Global land-ocean temperature index",
    caption = "Data source: NASA's Goddard Institute for Space Studies (GISS)"
  ) +
  # Set theme
  my_theme
```

## Temperature anomaly trend for each continent

```{r, echo=F, fig.width = 9, fig.height = 6}
# Kepp only year after 1880
df_continent_avg_temp_graph <- df_continent_avg_temp[df_continent_avg_temp$year > 1879, ]

ggplot(df_continent_avg_temp_graph ) +
  aes(x = year, y = lowess, colour = zone) +
  # Add horizontal and vertical line
  geom_hline(yintercept = 0, linetype="dashed", color = "grey") +
  geom_vline(xintercept = 1951, linetype="dashed", color = "grey") +
  geom_vline(xintercept = 1980, linetype="dashed", color = "grey") +
  # Plot line per continent
  geom_line(size = 0.7) +
  # Personalize legend
  guides(color = guide_legend(title = "")) +
  # Manual color 
  scale_color_manual(values = c(africa = "#151414", antarctica = "#BD9A00", asia = "#31B425",
                                europe = "#E90202", north_america = "#2C48E2", oceania = "#B280FC", 
                                south_america = "#9EE4F6"),
                     labels = c("Africa", "Antartica", "Asia", "Europe", "North America", "Oceania", "South America")) +
  # Add more ticks to x axis
  scale_x_continuous(breaks = round(seq(min(df_continent_avg_temp$year), 
                                        max(df_continent_avg_temp$year), by = 10), 1),
                     expand = c(0.007, 0)) + 
  # Add axes limits
  ylim(-1, 2) + 
  annotate("text", x=1995, y=1.6, label= "Europe shows the highest yearly 
average temperature anomaly", size = 2.5) +
    annotate("text", x=1965, y=-0.8, label= "Temperature anomaly are relative
to 1951-1980 average", size = 2.5) +
  # Add labels and text
  labs(
    x = "Year",
    y = "Temperature anomaly w.r.t 1951-80 (°C)",
    title = "Continent temperature anomaly",
    subtitle = "Global annual mean surface air temperature change",
    caption = "Data source: Berkley Earth."
  ) +
  # Set theme
  my_theme
```

## Temperature records months and regions

```{r}
# Order data by yearly anomaly
df_top_year <- df_continent_avg_temp[order(-df_continent_avg_temp$yearly_anomaly), ]
kable(head(df_top_year, 10))
```

```{r}
df_eu <- df_continent_avg_temp[df_continent_avg_temp$zone == 'europe', ]
write.csv(df_eu, "data/Berkley Earth/df_yearly_anomaly_EU.csv", row.names = TRUE)
```

